```{r 08_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# Interaction Models

## Learning Goals {-}

- Understand what research questions can be answered with interaction models
- Interpret coefficients in interaction models
- Connect visualizations to the coefficient estimates in interaction models
- Graphically/geometrically describe the "pictures" implied by different types of interaction models

<br><br><br><br>

## Discussion and Exercises {-}

A template RMarkdown document that you can start from is available [here](template_rmds/08-interaction-models.Rmd).

We will continue looking at the `diamonds` dataset.

- `carat`: the weight of the diamond in carats (1 carat = 200 milligrams)
- `price`: price in US dollars
- `cut`: quality of the cut of the diamond (Fair, Good, Very Good, Premium, Ideal)
- `color`: Level 1 (best) to Level 7 (worst)
- `clarity`: Level 1 (worst) to Level 8 (best)
- `x`, `y`, and `z`: length, width, and depth respectively (in mm)
- `depth`: total depth percentage = z / mean(x, y)
- `table`: width of top of diamond relative to widest point

```{r eval=TRUE, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)

diamonds <- read_csv("https://www.dropbox.com/s/9c8jqda4pwaq8i1/diamonds.csv?dl=1")
```

<br><br><br>

**Warm-up:** Consider the model of `price` as a function of `clarity` and `carat`.

$$ E[\hbox{price}] = \beta_0 + \beta_1\hbox{clarityLevel_2} + \beta_2\hbox{clarityLevel_3} + \cdots + \beta_7\hbox{clarityLevel_8} + \beta_8\hbox{carat} $$

- Why was this called a **parallel lines model**?
- Draw a picture of the relationships that this model implies and label the picture with the coefficients of the model.


<br><br><br>

**Exercise 1:** We looked at a scatterplot of `price` versus `carat` and boxplots of `price` versus `clarity`, but we haven't looked at all three variables simultaneously. Let's enrich the scatterplot to additionally show `clarity`.

We've seen that we can have `x`, `y`, `color`, and `fill` aesthetics (inside the `aes()`). How might we adapt the code below to show different color trend lines corresponding to the 8 clarity levels?

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
    geom_point() +
    geom_smooth()
```

<br><br><br>

If the relationship between the response $y$ and predictor $x_1$ differs depending on the value / categories of another predictor $x_2$, we say that $x_1$ and $x_2$ **interact**. We capture this in our models with an **interaction term**: $x_1 \times x_2$.

$$ E[y] = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_1 \times x_2 $$

The coefficient $\beta_3$ in front of the interaction term is called the **interaction coefficient**.

<br><br><br>

For now, let's work with a subset of the data that just has two clarity levels.

$$ E[\hbox{price}] = \beta_0 + \beta_1\hbox{clarity2} + \beta_2\hbox{carat} + \beta_3\hbox{clarity2}\times\hbox{carat} $$

```{r eval=TRUE}
diamonds_sub <- diamonds %>%
    filter(clarity %in% c("Level_1", "Level_2"))

mod1 <- lm(price ~ clarity*carat, data = diamonds_sub)
summary(mod1)
```

- **Note:** When you see a colon between two variables, the colon means a multiplication sign. (`clarityLevel_2:carat` is the product of the `clarityLevel_2` indicator variable and the `carat` variable.)

<br>

**Model formula**

$$
\begin{align*}
E[\hbox{price}] &= \beta_0 + \beta_1\hbox{clarity2} + \beta_2\hbox{carat} + \beta_3\hbox{clarity2}\times\hbox{carat} \\
&= -1480.56 - 1937.78\,\hbox{clarity2} + 4209.79\,\hbox{carat} + 3660.46\,\hbox{clarity2}\times\hbox{carat}
\end{align*}
$$

<br>

**Model formula for clarity level 1**

$$
\begin{align*}
E[\hbox{price}] &= -1480.56 - 1937.78\times 0 + 4209.79\,\hbox{carat} + 3660.46\times 0 \times\hbox{carat} \\
&= -1480.56 + 4209.79\,\hbox{carat}
\end{align*}
$$

<br>

**Model formula for clarity level 2**

$$
\begin{align*}
E[\hbox{price}] &= -1480.56 - 1937.78\times 1 + 4209.79\,\hbox{carat} + 3660.46\times 1 \times\hbox{carat} \\
&= (-1480.56 - 1937.78) + (4209.79 + 3660.46)\,\hbox{carat} \\
&= -3418.34 + 7870.25\,\hbox{carat}
\end{align*}
$$

<br>

**It is very helpful to draw a picture labeled with the coefficients.**

<br>

Coefficient interpretations:

- Intercept $\beta_0$: -1480.56    
    The average price for a zero carat, level 1 clarity diamond. (Could have made more meaningful by centering `carat`, but generally not interested in the intercept.)

- $\beta_2$ for `carat`: 4209.79    
    The slope for the clarity level 1 model. On average, diamond price increases by \$4209.79 per carat increase **in level 1 clarity diamonds**. (This rate of change is different for clarity level 2 diamonds.)

- $\beta_1$ for `clarityLevel_2`: -1937.78    
    The change in y-intercept going from clarity level 1 to 2.   
    On average, zero carat diamonds that are clarity level 2 are worth \$1937.78 less than zero carat, clarity level 1 diamonds.

- $\beta_3$ for the interaction term (`clarityLevel_2:carat`): 3660.46    
    The change in slope for the clarity level 2 vs. the clarity level 1 model. The average change in price per carat increase is 3660.46 ($/carat) higher for clarity level 2 vs. clarity level 1 diamonds.    
    The slope is steeper for level 2. A carat is worth more for a level 2 clarity diamond.


<br><br><br>

**Exercise 2:** Fit a model with interaction between `clarity` and `carat` for the full diamonds dataset.

a. Write out the model formulas for clarity level 1, 2, ... diamonds until you get the hang of it.

b. Draw a picture with the model lines for clarity level 1, 2, ... 8 diamonds. Annotate this picture with the model coefficients.

c. Interpret the `clarityLevel_3` and `clarityLevel_3:carat` coefficients. Repeat for level 4 and onwards until you get the hang of it.

d. For which clarity level is a carat worth the most? How much is a carat worth on average?

<br><br><br>

**Exercise 3:** Now let's consider diamond color (Level 1 is best and Level 7 is worst).

a. What research question are you exploring when you fit a model with interaction between `color` and `carat`? Generalize this to a broader context.

b. Make a plot that allows you to explore potential interaction between `color` and `carat`. What are you looking to see in the plot? Do you think an interaction model is appropriate?

c. Fit the interaction model, and focus on the coefficients of substantive interest. Summarize your overall conclusions.



