```{r 13_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, fig.align="center", fig.width=8, fig.height=4)
```

# Causal Inference

## Learning Goals {-}

- Connect the idea of a causal effect to the idea of confounding
- Use causal diagrams to identify which variables should be "controlled"/"adjusted" for in analyses
- Use regression models to answer causal questions


<br><br><br><br>


## Discussion {-}

> "Correlation does not imply causation."

Your instructor finds this saying vacuous and "avoidist".

- "Avoidist": cultural tendency to (unreasonably) avoid talking about a topic
- Why do people always jump to this saying?


<br><br><br><br>


- Shoe size and reading ability are positively correlated
- Ice cream sales and drownings
- Can you think of others?
- **Question:** What causal diagrams are behind these situations?


<br><br><br><br>


```{r fig.width=12, echo=FALSE, eval=TRUE}
par(mar = rep(1,4), mfrow = c(1,2))
## Shoe size and reading ability
plot(1, type = "n", xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", xlim = c(0,6), ylim = c(1,4))
text(c("Shoe size", "Age", "Reading ability"), x = c(1,3,5), y = c(1,4,1), cex = 1.1)
arrows(x0 = c(3,3), y0 = c(4,4)-0.2, x1 = c(1,5), y1 = c(1,1)+0.2, angle = 25, lwd = 4)
## Ice cream sales and drownings
plot(1, type = "n", xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", xlim = c(0,6), ylim = c(1,4))
text(c("Ice cream sales", "Temperature", "Drownings"), x = c(1,3,5), y = c(1,4,1), cex = 1.1)
arrows(x0 = c(3,3), y0 = c(4,4)-0.2, x1 = c(1,5), y1 = c(1,1)+0.2, angle = 25, lwd = 4)
```

- Holding age constant would show zero effect of shoe size on reading ability
- Holding temperature constant would show zero effect of ice cream sales on drownings


<br><br><br><br>


In most research settings, there is an arrow between the variable of interest and the response and many potential confounding variables:

```{r echo=FALSE, eval=TRUE}
plot(1, type = "n", xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", xlim = c(0,6), ylim = c(1,4))
text(c("X", "C1, C2, C3, ...", "Y"), x = c(1,3,5), y = c(1,4,1), cex = 1.1)
arrows(x0 = c(3,3), y0 = c(4,4)-0.2, x1 = c(1,5), y1 = c(1,1)+0.2, angle = 25, lwd = 4)
arrows(x0 = 1.7, y0 = 1, x1 = 4.5, y1 = 1, angle = 25, lwd = 4)
```

- The goal of causal inference is **quantitatively estimating** the effect of $X$ on $Y$ along the direct arrow.
- The "isolated effect" that we've been referring to is actually another way of phrasing the **causal effect**.
- If we can measure all confounders, including all confounders in a regression model allows us to hold those variables constant.
    - Also called "controlling for" confounders
    - Also called "adjusting for" confounders

<br>

In a non-interaction model, $\beta_1$ is of interest (`Y ~ X + C1 + C2`):

$$ E[Y] = \beta_0 + \beta_1\,X + \beta_2\,C_1 + \beta_3\,C_2 $$


An interaction model also allows us to hold the confounders constant (`Y ~ X*C1 + C2`):

$$ E[Y] = \beta_0 + \beta_1\,X + \beta_2\,C_1 + \beta_3\,X\times C_1 + \beta_4\,C_2 $$

Or `Y ~ X*C2 + C1`:

$$ E[Y] = \beta_0 + \beta_1\,X + \beta_2\,C_2 + \beta_3\,X\times C_2 + \beta_4\,C_1 $$

(You'll practice in the Exercises.)


<br><br><br><br>


**Overadjustment bias**

It is tempting to put as many variables as we can into models to hold constant any potential confounders.

- It is very easy to do harm here!
- It is possible to *overadjust*.

<br><br>

- To understand the causal effect of the flu shot, we should adjust for (hold constant) age and socioeconomic status.
- Do we really want to adjust for immune cell level in the time period after the shot?
    - Immune cell levels are a **mediator** of the effect of the flu shot on subsequent flu.
    - This variable lies on a **causal pathway** between the variable of interest and the response.

```{r echo=FALSE, eval=TRUE}
par(mfrow = c(1,1))
plot(1, type = "n", xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", xlim = c(0,10), ylim = c(1,4))
text(c("Flu shot", "Age, SES", "Flu"), x = c(1,5,9), y = c(1,4,1), cex = 1.1)
arrows(x0 = c(5,5), y0 = c(4,4)-0.2, x1 = c(1,9), y1 = c(1,1)+0.2, angle = 25, lwd = 4)
text("Immune cells", x = 5, y = 1, cex = 1.1)
arrows(x0 = 1.7, y0 = 1, x1 = 3.5, y1 = 1, angle = 25, lwd = 4)
arrows(x0 = 6.5, y0 = 1, x1 = 8.5, y1 = 1, angle = 25, lwd = 4)
```


<br><br><br><br>


**Would we ever want to adjust for mediators?**

```{r echo=FALSE, eval=TRUE}
par(mfrow = c(1,1))
plot(1, type = "n", xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", xlim = c(0,10), ylim = c(1,4))
text(c("Flu shot", "Age, SES", "Flu"), x = c(1,5,9), y = c(1,4,1), cex = 1.1)
arrows(x0 = c(5,5), y0 = c(4,4)-0.2, x1 = c(1,9), y1 = c(1,1)+0.2, angle = 25, lwd = 4)
text("Immune cells", x = 5, y = 1, cex = 1.1)
arrows(x0 = 1.7, y0 = 1, x1 = 3.5, y1 = 1, angle = 25, lwd = 4)
arrows(x0 = 6.5, y0 = 1, x1 = 8.5, y1 = 1, angle = 25, lwd = 4)
text("Hand washing", x = 5, y = 1.5, cex = 1.1)
arrows(x0 = 1.7, y0 = 1.1, x1 = 3.5, y1 = 1.5, angle = 25, lwd = 4)
arrows(x0 = 6.5, y0 = 1.5, x1 = 8.5, y1 = 1.2, angle = 25, lwd = 4)
```

- The **total effect** of the flu shot on subsequent flu is the effect through *both* the immune cells and hand washing pathways.
    - If we adjust only for age and SES (hold them constant), we'll estimate the total effect.
- The **direct effect** (not pictured) would be the effect along an arrow directly from `Flu shot` to `Flu`.
- We are interested in the **indirect effect** through `Immune cells`.
    - If we adjust for age, SES, and hand-washing, we can estimate this indirect effect.


<br><br><br><br>


Instead of...

> "Correlation does not imply causation."


...let's say

> "Properly adjusted correlations are causation."


<br><br><br><br>


## Exercises {-}

You won't need R for these conceptual exercises.

### Exercise 1 {-}

Draw a causal diagram corresponding to a study of how yoga participation affects stress levels. What confounding variables are likely present? What possible mediating variables would we *want to* adjust for in our analysis? What mediating variables would *not want to* adjust for?

<br><br>

### Exercise 2 {-}

Suppose that in the yoga study, we conclude that we need to adjust for the confounding variables of age and prior lifetime history of exercise. The variables recorded are:

- `yoga`: yes/no participation
- `stress`: hours per week under stress
- `age`: in years
- `exercise`: in hours per week

a. Write out the multiple regression model (without interaction) that could be used to estimate the causal effect of yoga on stress levels.

b. Is there a real, meaningful effect of yoga on stress levels? Describe in detail how we could answer the question.

c. Suppose you made a plot that indicates an interaction between `yoga` and `age` in predicting stress levels (`age` and `exercise` are still confounders). What would this plot look like?

d. Continuing from (c), consider the research question: "How does the causal effect of yoga on stress differ across ages?"    
    - Write down the regression model formula that would allow you to answer this research question.
    - There isn't necessarily a single coefficient from this model that will answer the question, but describe how you could use the model to make predictions that would answer this research question.

<br><br>

### Exercise 3 {-}

Suppose now that `yoga` and `age` are now measured differently:

- `yoga`: hours per week of participation
- `age`: two categories "young" and "old"

a. Suppose you made a plot that indicates an interaction between `yoga` and `age` in predicting stress levels (`age` and `exercise` are still confounders). What would this plot look like?

b. Continuing from (a), consider the research question: "How does the causal effect of yoga on stress differ across ages?"    
    - Write down the regression model formula that would allow you to answer this research question.
    - There are two coefficients in this model that are of interest. What are they, and how can we interpret them in a contextually meaningful way?
    - Describe how we could assess if there is a real, meaningful difference in effects of yoga between the young and old.


