```{r 16_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=TRUE)
```

# Simple Logistic Regression

## Learning Goals {-}

- Construct simple logistic regression models in R
- Interpret coefficients in simple logistic regression models
- Use simple logistic regression models to make predictions
- Describe the form (shape) of relationships on the log odds, odds, and probability scales

<br><br><br><br>

## Warm-up {-}

Navigate to:

<center>
[PollEv.com/lesliemyint417](https://www.PollEv.com/lesliemyint417)
</center>

<br><br><br><br>

## Exercises {-}

A template RMarkdown document that you can start from is available [here](template_rmds/16-simple-logistic-regression.Rmd).

We'll look at the O-ring data described in the video. Load the data and necessary packages as follows:

```{r}
library(readr)
library(dplyr)
library(ggplot2)

oring <- read_csv("https://www.macalester.edu/~ajohns24/data/NASA.csv")
```

Let's get acquainted with the data and make a few exploratory plots:

```{r}
dim(oring)
head(oring)

# Univariate visualization of Broken


# Univariate visualization of Temp


# Visualization of Broken and Temp
```

### Exercise 1 {-}

a. We want to model `Broken` as a function of `Temp`. Write down the logistic regression model formula. Try to do these without referring to notes.

b. We can fit logistic regression models in R using the `glm()` function. The "lm" part of "glm" stands for "linear model" (just like the `lm()` function), and the "g" stands for "generalized". (The left hand side of the model has been made more general than just $E[Y]$.) Also note that we need to supply the argument `family = "binomial"`.
    ```{r}
    oring_mod <- glm(Broken ~ Temp, data = oring, family = "binomial")
    summary(oring_mod)
    ```
    We'll focus primarily on the output in the coefficients table. Interpret the intercept and the temperature coefficient on the log scale in a contextually meaningful way. What concern arises in interpreting the intercept?

c. Interpret the coefficients *on the natural scale* by exponentiating. Use the `exp()` function to do this. (e.g., `exp(3)` gives $e^3$.)

d. Use this model to make predictions about a 60 degree day:
    - Predict the log odds of O-ring failure
    - Predict the odds of O-ring failure
    - Predict the probability of O-ring failure. Also write this probability using conditional probability notation.

<br><br>

### Exercise 2 {-}

a. When you use `glm()`, R by default will use `Broken = 1` as the event of interest (as opposed to `Broken = 0`). Suppose that we were to fit the logistic regression model where `Broken = 0` became the event of interest. Mathematically work out what the new values of the model coefficients will be.

b. Check your work by replacing the `glm()` model formula with `Broken==0 ~ Temp`.

<br><br>

### Exercise 3 {-}

Let's visualize the model's predictions.

a. The code below adds the predicted log odds to the original `oring` dataset. Complete the code to also add the predicted odds and probabilities.
    ```{r}
    oring <- oring %>%
        mutate(
            log_odds = predict(oring_mod),
            odds = ???, # Hint: you can use log_odds in this expression
            prob = ???
        )
    ```

b. Construct plots of the log odds, odds, and probability as a function of temperature.

c. We can also zoom out to see a broader temperature range. Describe the shape of the relationship between the probability of an O-ring breaking and temperature.
    ```{r}
    ggplot(oring, aes(x = Temp, y = Broken)) +
        geom_point() +
        geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE, fullrange = TRUE) + 
        labs(y = "probability of breaking") + 
        lims(x = c(0,100))
    ```
    Note: the function shown in this plot is called the **logistic function** which is where logistic regression gets its name.

<br><br>

### Exercise 4 {-}

We have the following data on whether or not individuals were exposed to an environmental hazard and whether or not they developed a certain disease within 10 years.

|           | Disease | No disease | Total |
|-----------|---------|------------|-------|
| Exposed   | 5       | 495        | 500   |
| Unexposed | 5       | 995        | 1000  |
| Total     | 10      | 1490       | 1500  |

Write down the logistic regression model formula corresponding and estimate the coefficients by hand using data from the table. (Use a variable called `exposed` that is equal to 1 if exposed and 0 if not.)

