```{r 17_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=TRUE)
```

# Multiple Logistic Regression

## Learning Goals {-}

- Construct multiple logistic regression models in R
- Construct and interpret visualizations to inform model building
- Interpret multiple logistic regression model output

<br><br><br><br>

## Exercises {-}

A template RMarkdown document that you can start from is available [here](template_rmds/17-multiple-logistic-regression.Rmd).

We will be looking at data from the Add Health study (https://cpc.unc.edu/projects/addhealth), a national survey of adolescent health. 

> **Research question:** What is the causal effect of adolescent marijuana smoking on adulthood cigarette smoking?

The following variables are in the dataset:

- `exposure`: Indicator for whether or not marijuana was smoked in adolescence. 1=yes, 0=no. (Main predictor of interest.)
- `smoke`: Indicator for whether the individual smoked in adulthood. 1=yes, 0=no. (Response variable.)
- `age`: Age at study entry (continuous)
- `male`: Indicator for whether the individual was male. 1=yes, 0=no.
- `white`: Indicator for whether the individual was White. 1=yes, 0=no.
- `susp`: Indicator for whether the individual was ever suspended from school. 1=yes, 0=no.
- `mathG`: Math grade (1 to 4 corresponding respectively to A to D)
- `readG`: Reading grade (1 to 4 corresponding respectively to A to D)
- `parentED`: Parent education (1 to 5 corresponding respectively to less than high school, high school, vocational school, some college, and college graduate and beyond).
- `housesmoke`: Indicator for the presence of a cigarette smoker in the individual's household. 1=yes, 0=no.

```{r}
library(readr)
library(ggplot2)
library(dplyr)

addhealth <- read_csv("https://www.dropbox.com/s/h09nq0nwgx8eioz/addhealth.csv?dl=1")
```

Peek at the first few rows:

```{r}
head(addhealth)
```

**Note:** Throughout these exercises, you'll want to use `factor()` within visualizations and model formulas to force R to treat a variable as categorical (rather than quantitative).

<br><br>

### Exercise 1 {-}

Let's get a feel for the main variables of interest.

Construct and interpret appropriate univariate visualizations of adolescent marijuana smoking and adulthood cigarette smoking.

<br><br>

### Exercise 2 {-}

a. Construct and interpret an appropriate visualization of the relationship between `exposure` and `smoke`.
b. Based on your visualization what would you expect about the coefficient in an appropriate model of `smoke ~ exposure`? What would you expect about the 95% bootstrap confidence interval?
c. Fit the appropriate model, and interpret the coefficient of interest on the natural scale.
d. Construct a 95% bootstrap confidence interval for the coefficient (on the log scale). Transform the confidence interval to the natural scale by exponentiating both endpoints. If we define a meaningful effect as one that has an odds ratio of at least 1.5, do we have evidence for a meaningful effect of adolescent marijuana smoking in the broader population?

<br><br>

### Exercise 3 {-}

That previous analysis ignored the "causal" part of the question!

a. Looking through the variables measured, what variables might be confounders and what might be mediators?
b. Make appropriate plots to determine whether `housesmoke` might confound the relationship between adolescent marijuana smoking and adulthood cigarette smoking.
c. Based on your results from part (b), fit an appropriate model, and interpret the coefficient of interest on the natural scale.
d. Construct a 95% bootstrap confidence interval for the coefficient on the natural scale. Interpret the interval (what information does it contain?), and use it to determine whether we have evidence for a real, meaningful causal effect of adolescent marijuana smoking in the broader population.

