```{r 21_setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# Hypothesis Testing

## Learning Goals {-}

- Explain the logic of hypothesis testing
- Conduct hypothesis tests in the context of regression models to answer research questions
- Relate hypothesis testing to confidence interval procedures
- Interpret quantities related to hypothesis testing

<br><br><br><br>

## Discussion {-}

When we fit regression models in R and display the `summary()` output, the Coefficients table gives two new quantities used in hypothesis testing. By default, a null value of zero is used (i.e., $H_0: \beta=0$).

For linear regression models:

- `t value`: the test statistic called "t" because the sampling distribution of these test statistics follows a **t distribution** (very closely related to the normal distribution)
- `Pr(>|t|)`: shorthand for **pr**obability that the test statistic is larger in absolute value than the observed test statistic $t$ ($|t|$ represents the absolute value of $t$). This is the **two-sided p-value** for the test.

```
> lm(Price ~ Age, data = homes) %>% summary()

Coefficients:
             Estimate Std. Error t value Pr(>|t|)    
(Intercept) 229728.46    3218.18  71.385  < 2e-16 ***
Age               ???      79.66  -7.987  2.5e-15 ***
```

**Question:** The estimate for `Age` was deleted by accident! Are we able to fill in that missing information? Why or why not?

<br><br>

For logistic regression models, the "t" is instead a "z" because the sampling distribution of coefficients is the normal distribution. (Usually quantities that follow a normal distribution are denoted by "z".)

```
> glm(smoke ~ exposure, data = addhealth, family = "binomial") %>% summary()

Coefficients:
            Estimate Std. Error z value Pr(>|z|)    
(Intercept) -0.86275    0.05805 -14.862  < 2e-16 ***
exposure     0.91318    0.19237   4.747 2.06e-06 ***
```

<br><br>

We noted in the video that both confidence intervals and hypothesis tests can answer questions about whether we have evidence for real, meaningful broader population effects.

It turns out that decisions made from hypothesis tests with a significance level of $\alpha$ are equivalent to the corresponding $100(1-\alpha)\%$ CI.

- e.g., Conducting a hypothesis test with a significance level of $\alpha = 0.05$ is equivalent to conclusion from a 95% CI


<br><br><br><br>


## Exercises {-}

A template RMarkdown document that you can start from is available [here](template_rmds/21-hypothesis-testing.Rmd).

We will be using data on 1167 colleges obtained from an edition of the U.S. News College Rankings. For each college, the following variables are given:

- `College`: Name of school
- `GradRate`: College graduation rate (as a value from 0 to 100)
- `SFRatio`: Student-to-faculty ratio
- `AdmisRate`: Percentage of applicants accepted (as a value from 0 to 100)
- `FacultyPhD`: Percentage of faculty with a PhD (as a value from 0 to 100)
- `Type`: School type (Private or Public)
- `Region`: Location of school: Midwest, NorthEast, South, or West.
- `State`: School state
- `MathSAT`: Average Math SAT score of entering first-years
- `VerbalSAT`: Average Verbal SAT score of entering first-years
- `ACT`: Average ACT score of entering first-years

```{r}
library(readr)
library(dplyr)
library(ggplot2)

colleges <- read_csv("https://www.dropbox.com/s/efgkjbqeiikzzn7/USNews.csv?dl=1")
```

> **Research question:** What is the causal effect of student-to-faculty ratios on graduation rates? How does this effect differ across school types?

Throughout, use zero as the criteria for a meaningful difference and a significance level of 0.01.

### Exercise 1 {-}

a. Fit an *unadjusted* model that ignores any potential confounders. Present a (naïve) answer to the first research question by interpreting the relevant coefficient and conducting an appropriate hypothesis test.

b. What would you expect about the corresponding 99% confidence interval (CI) for the coefficient (whether through bootstrapping or the CLT)? What about the 95% CI for the coefficient? Briefly explain, and check your answer with `confint()`.

<br><br>

### Exercise 2 {-}

We find that school type (`Type`) and average ACT scores (`ACT`) are potential confounders of the relationship between student-to-faculty ratios and graduation rates.

a. Fit an *adjusted* model that answers the first research question. **Give an interpretation** for all 4 quantities for the relevant coefficient: the estimate, the standard error, the test statistic, and p-value.

b. Show that you reach the same conclusions with an appropriate confidence interval.

<br><br>

### Exercise 3 {-}

Fit an appropriate model to answer the second research question. Support your response by interpreting the two relevant coefficients and by appropriately using a p-value and a confidence interval.

<br><br>

### Exercise 4 {-}

Suppose we know that the distribution of a test statistic $Z$ "under $H_0$" (if $H_0$ is true) is the standard normal distribution (mean = 0, SD = 1). This is the case in logistic regression.

a. For what values of the test statistic $Z$ will I reject $H_0$ using $\alpha = 0.05$ as my p-value cutoff?

b. For what values of the test statistic $Z$ will I reject $H_0$ using $\alpha = 0.32$ as my p-value cutoff?
