---
title: "Multiple Linear Regression"
author: "Your Name"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Setup

```{r}
library(readr)
library(ggplot2)
library(dplyr)

diamonds <- read_csv("https://www.dropbox.com/s/9c8jqda4pwaq8i1/diamonds.csv?dl=1")
```

**Research question:** How are different factors related to the price of a diamond?

Let's familiarize ourselves with the data:

```{r}
# Display the dimensions and first few rows
```

Key variables:

- `carat`: the weight of the diamond in carats (1 carat = 200 milligrams)
- `price`: price in US dollars
- `cut`: quality of the cut of the diamond (Fair, Good, Very Good, Premium, Ideal)
- `color`: Level 1 (worst) to Level 7 (best)
- `clarity`: Level 1 (worst) to Level 8 (best)
- `x`, `y`, and `z`: length, width, and depth respectively (in mm)
- `depth`: total depth percentage = z / mean(x, y)
- `table`: width of top of diamond relative to widest point


# Exercises

## Exercise 1

What are your expectations for how `price` and `clarity` will be related? Construct a visualization to see. (Clarity level 1 is deemed worst and level 8 is deemed best.)

```{r}
# Plot of price vs. clarity
```






## Exercise 2

Let's first fit a linear regression model with just `clarity` as a predictor of `price`.

```{r}
mod1 <- lm(price ~ clarity, data = diamonds)
summary(mod1)
```

a. Why is there no coefficient for `clarityLevel_1`?

b. Interpret the `clarityLevel_2` and `clarityLevel_5` coefficients.

c. Based on this output, can you tell which clarity level has the highest mean price in this dataset? Explain.

d. Relate the magnitudes of the coefficients to the trends that you see in your plot.






## Exercise 3

What if the most clear diamonds happened to be quite small (low carat)? Would this explain why your plot looks the way it does?







## Exercise 4

Add to your previous model by including `carat` as a predictor.

```{r}
mod2 <- lm(price ~ clarity+carat, data = diamonds)
summary(mod2)
```

a. Write an expression for the average price of a clarity level 1 diamond that is zero carats. What coefficient does this allow you to interpret?

b. Write an expression for the average price of a clarity level 1 diamond that is $C$ carats.

c. Write an expression for the average price of a clarity level 2 diamond that is $C$ carats. Subtract your answer from part (b) from this expression. Based on this, give an interpretation of the `clarityLevel_2` coefficient.

d. Write an expression for the average price of a clarity level 3 diamond that is $C$ carats. Subtract your answer from part (b) from this expression. Based on this, give an interpretation of the `clarityLevel_3` coefficient.

e. Pick any clarity level. Write an expression for the average price of two diamonds of that clarity level: (1) one with $C$ carats and (2) one with $C+1$ carats. Subtract (1) from (2), and thus interpret the `carat` coefficient.

f. Compare the values of the `clarityLevel` coefficients in `mod2` and `mod1` and form a conclusion about the relationship between `price` and `clarity`.






