---
title: "Interaction Models"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Setup

```{r}
library(readr)
library(ggplot2)
library(dplyr)

diamonds <- read_csv("https://www.dropbox.com/s/9c8jqda4pwaq8i1/diamonds.csv?dl=1")
```

- `carat`: the weight of the diamond in carats (1 carat = 200 milligrams)
- `price`: price in US dollars
- `cut`: quality of the cut of the diamond (Fair, Good, Very Good, Premium, Ideal)
- `color`: Level 1 (best) to Level 7 (worst)
- `clarity`: Level 1 (worst) to Level 8 (best)
- `x`, `y`, and `z`: length, width, and depth respectively (in mm)
- `depth`: total depth percentage = z / mean(x, y)
- `table`: width of top of diamond relative to widest point

# Discussion and Exercises

## Exercise 1

We looked at a scatterplot of `price` versus `carat` and boxplots of `price` versus `clarity`, but we haven't looked at all three variables simultaneously. Let's enrich the scatterplot to additionally show `clarity`.

We've seen that we can have `x`, `y`, `color`, and `fill` aesthetics (inside the `aes()`). How might we adapt the code below to show different color trend lines corresponding to the 8 clarity levels?

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
    geom_point() +
    geom_smooth()
```



## Exercise 2

Fit a model with interaction between `clarity` and `carat` for the full diamonds dataset.

a. Write out the model formulas for clarity level 1, 2, ... diamonds until you get the hang of it.

b. Draw a picture with the model lines for clarity level 1, 2, ... 8 diamonds. Annotate this picture with the model coefficients.

c. Interpret the `clarityLevel_3` and `clarityLevel_3:carat` coefficients. Repeat for level 4 and onwards until you get the hang of it.

d. For which clarity level is a carat worth the most? How much is a carat worth on average?






## Exercise 3

Now let's consider diamond color (Level 1 is best and Level 7 is worst).

a. What research question are you exploring when you fit a model with interaction between `color` and `carat`? Generalize this to a broader context.

b. Make a plot that allows you to explore potential interaction between `color` and `carat`. What are you looking to see in the plot? Do you think an interaction model is appropriate?

c. Fit the interaction model, and focus on the coefficients of substantive interest. Summarize your overall conclusions.

















