---
title: "Multiple Logistic Regression"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Setup

> **Research question:** What is the causal effect of adolescent marijuana smoking on adulthood cigarette smoking?

The following variables are in the dataset:

- `exposure`: Indicator for whether or not marijuana was smoked in adolescence. 1=yes, 0=no. (Main predictor of interest.)
- `smoke`: Indicator for whether the individual smoked in adulthood. 1=yes, 0=no. (Response variable.)
- `age`: Age at study entry (continuous)
- `male`: Indicator for whether the individual was male. 1=yes, 0=no.
- `white`: Indicator for whether the individual was White. 1=yes, 0=no.
- `susp`: Indicator for whether the individual was ever suspended from school. 1=yes, 0=no.
- `mathG`: Math grade (1 to 4 corresponding respectively to A to D)
- `readG`: Reading grade (1 to 4 corresponding respectively to A to D)
- `parentED`: Parent education (1 to 5 corresponding respectively to less than high school, high school, vocational school, some college, and college graduate and beyond).
- `housesmoke`: Indicator for the presence of a cigarette smoker in the individual's household. 1=yes, 0=no.

```{r}
library(readr)
library(ggplot2)
library(dplyr)

addhealth <- read_csv("https://www.dropbox.com/s/h09nq0nwgx8eioz/addhealth.csv?dl=1")
```

Peek at the first few rows:

```{r}
head(addhealth)
```

**Note:** Throughout these exercises, you'll want to use `factor()` within visualizations and model formulas to force R to treat a variable as categorical (rather than quantitative).

<br><br>

## Exercise 1

Let's get a feel for the main variables of interest.

Construct and interpret appropriate univariate visualizations of adolescent marijuana smoking and adulthood cigarette smoking.



<br><br>



## Exercise 2

### Part a

Construct and interpret an appropriate visualization of the relationship between `exposure` and `smoke`.

### Part b

Based on your visualization what would you expect about the coefficient in an appropriate model of `smoke ~ exposure`? What would you expect about the 95% bootstrap confidence interval?

### Part c

Fit the appropriate model, and interpret the coefficient of interest on the natural scale.

### Part d

Construct a 95% bootstrap confidence interval for the coefficient (on the log scale). Transform the confidence interval to the natural scale by exponentiating both endpoints. If we define a meaningful effect as one that has an odds ratio of at least 1.5, do we have evidence for a meaningful effect of adolescent marijuana smoking in the broader population?



<br><br>



## Exercise 3

That previous analysis ignored the "causal" part of the question!

### Part a

Looking through the variables measured, what variables might be confounders and what might be mediators?

### Part b

Make appropriate plots to determine whether `housesmoke` might confound the relationship between adolescent marijuana smoking and adulthood cigarette smoking.

### Part c

Based on your results from part (b), fit an appropriate model, and interpret the coefficient of interest on the natural scale.

### Part d

Construct a 95% bootstrap confidence interval for the coefficient on the natural scale. Interpret the interval (what information does it contain?), and use it to determine whether we have evidence for a real, meaningful causal effect of adolescent marijuana smoking in the broader population.


